---
title: "P8105 Homework 1"
author: "Ruiyang Li"
output: github_document
---

This is my solution to HW1. 

```{r setup}
library(tidyverse)
```


# Problem 1

Create a data frame with the specified elements. 

```{r create_df}
## set seed to ensure reproducibility
set.seed(20200919)

## create data frame
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_pos = samp > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vec_factor = factor(c("low", "low", "low", "mod", "mod", "mod", "high", "high", "high", "high"))
  )
```

Take the mean of each variable in my data frame.

```{r take_mean}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_pos))
mean(pull(prob1_df, vec_char))
mean(pull(prob1_df, vec_factor))
```

Observations:

* R computes the mean of the numerical variable `samp` to be `r round(mean(pull(prob1_df, samp)), digits = 2)`.
* R computes the mean of the logical variable `samp_pos` to be `r round(mean(pull(prob1_df, samp_pos)), digits = 2)`.
* However, R fails to compute the mean for the character variable `vec_char` and the factor variable `vec_factor`. 

Therefore, we can take the mean of numbers and logical but not character or factor. 


Apply the `as.numeric` function to the logical, character, and factor variables.

```{r convert_to_numeric, eval = FALSE}
as.numeric(pull(prob1_df, samp_pos))
as.numeric(pull(prob1_df, vec_char))
as.numeric(pull(prob1_df, vec_factor))
```

What happens in the above code chunk is that we can convert the logical variable and the factor variable to numbers using the `as.numeric` function, but we fail to do so for the character variable. 

The reason why this happens is that 

* for the logical variable, the `as.numeric` function converts those with "FALSE" values to 0 and those with "TRUE" values to 1
* for the factor variable, the `as.numeric` function converts each of the 3 levels to 1, 2, and 3
* however, for the character variable, NAs are introduced by coercion by the `as.numeric` function

This helps explain what happens when we try to take the mean: 

* we know that we can only compute mean on numerical data
* the mean of the numerical variable `samp` can be computed directly
* the underlying way how R computes the mean of the logical variable is that it first converts the logical variable into the numerical variable and then uses the converted numerical variable to compute the mean 
* as expected, R fails to compute the mean for character variables and factor variables


In the next code chunk, we 

* convert the logical vector to numeric, and multiply the random sample by the result
* convert the logical vector to a factor, and multiply the random sample by the result
* convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r convert_variables}
as.numeric(pull(prob1_df, samp_pos)) * pull(prob1_df, samp)
as.factor(pull(prob1_df, samp_pos)) * pull(prob1_df, samp)
as.numeric(as.factor(pull(prob1_df, samp_pos))) * pull(prob1_df, samp)
```

