---
title: "P8105 HW1"
author: "Ruiyang Li"
output: github_document
---

This is my solution to HW1. 

```{r setup}
library(tidyverse)
```


# Problem 1

Create a data frame with the specified elements. 

```{r create_df}
## set seed to ensure reproducibility
set.seed(20200919)

## create data frame
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_pos = samp > 0,
    vec_char = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    vec_factor = factor(c("low", "low", "low", "mod", "mod", "mod", "high", "high", "high", "high"))
  )
```

Take the mean of each variable in my data frame.

```{r}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_pos))
mean(pull(prob1_df, vec_char))
mean(pull(prob1_df, vec_factor))
```

Observations:

* R computes the mean of the numerical variable `samp` to be `r round(mean(pull(prob1_df, samp)), digits = 2)`.
* R computes the mean of the logical variable `samp_pos` to be `r round(mean(pull(prob1_df, samp_pos)), digits = 2)`.
* However, R fails to compute the mean for the character variable `vec_char` and the factor variable `vec_factor`. 


